---
title: "Idealized Forms in Early Concepts of the Biological World: Results"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#Load libraries needed for these analyses (any uninstalled packages will need to be installed)

library(tidyverse)
library(modelr)
library(geepack)
library(doBy)
library(ggplot2)
library(ggthemes)
library(visreg)
library(effects)
library(dplyr)
library(forcats)
library(lme4)
library(effects)
library(devtools)
library(papaja)
library(multcomp)
library(aod)
library(knitr)
library(car)
library(RColorBrewer)
library(cowplot)
library(lmerTest)
```

```{r, include = FALSE}
##### Study 1 #####

#Study 1
#Read in the data file
#This file contains data from children and adults, in wide format
ex1Wide<- read_csv("Extremes1_Items.csv")

#Turn categorical variables into factors, and set reference categories and labels in a sensible and useful way

ex1Wide$Condition<-factor(ex1Wide$Condition, levels = c("1", "2"))
print(levels(ex1Wide$Condition))
ex1Wide$Condition<-recode_factor(ex1Wide$Condition, "1" = "Best", "2" = "Representative")

ex1Wide$Gender<-factor(ex1Wide$Gender, levels = c("1", "2"))
print(levels(ex1Wide$Gender))
ex1Wide$Gender<-recode_factor(ex1Wide$Gender, "1" = "Male", "2" = "Female", "3" = "Other")
print(levels(ex1Wide$Gender))

ex1Wide$AgeGroup<-factor(ex1Wide$AgeGroup, levels = c("1", "2", "3"))
print(levels(ex1Wide$AgeGroup))
ex1Wide$AgeGroup<-recode_factor(ex1Wide$AgeGroup, "1" = "Child_5.6", "2" = "Child_7.8", "3" = "Adult")

ex1Wide$AdultChild <- ex1Wide$AgeGroup
ex1Wide$AdultChild<-recode_factor(ex1Wide$AdultChild, "Child_5.6" = "Child", "Child_7.8" = "Child","Adult" = "Adult")

ex1Wide_Child <- ex1Wide %>%
  filter(AdultChild == "Child")
ex1Wide_Adult <- ex1Wide %>%
  filter(AdultChild == "Adult")
ex1Wide_Child_5.6 <- ex1Wide_Child %>%
  filter(AgeGroup == "Child_5.6")
ex1Wide_Child_7.8 <- ex1Wide_Child %>%
  filter(AgeGroup == "Child_7.8")

#Nature and science questionnaire info - make Community a factor to specify levels
ex1Wide$Community<-factor(ex1Wide$Community, levels = c("0", "1", "2"))
print(levels(ex1Wide$Community))
ex1Wide$Community<-recode_factor(ex1Wide$Community, "0" = "Urban", "1" = "Suburban", "2" = "Rural")
print(levels(ex1Wide$Community))

#Tidy the data
#For this dataset, that means making two separate tidy columns - one for choice (representative or ideal) and one for induction

#Tidy - choice items
#we will add the discover items in a separate column after we tidy the choice into a column
ex1TidyT <- ex1Wide %>%
  gather(item, choice, choice1, choice2, choice3, choice4, choice5, choice6, choice7, choice8)

ex1TidyT$item<-factor(ex1TidyT$item, levels = c("choice1", "choice2", "choice3", "choice4", "choice5", "choice6", "choice7", "choice8"))
print(levels(ex1TidyT$item))
ex1TidyT$item<-recode_factor(ex1TidyT$item, "choice1" = "1", "choice2" = "2", "choice3" = "3", "choice4" = "4", "choice5" = "5", "choice6" = "6", "choice7" = "7", "choice8" = "8")

keepvarsT <- c("SID", "Condition", "ExactAge", "AgeGroup", "item", "choice")
ex1TidyT <- ex1TidyT[keepvarsT]

#Tidy - induction items
ex1TidyD <- ex1Wide %>%
  gather(item, induction, induction1, induction2, induction3, induction4, induction5, induction6, induction7, induction8)

keepvarsD <- c("SID", "item", "induction")
ex1TidyD_short <- ex1TidyD[keepvarsD]

ex1TidyD_short$item<-factor(ex1TidyD_short$item, levels = c("induction1", "induction2", "induction3", "induction4", "induction5", "induction6", "induction7", "induction8"))
print(levels(ex1TidyD_short$item))
ex1TidyD_short$item<-recode_factor(ex1TidyD_short$item, "induction1" = "1", "induction2" = "2", "induction3" = "3", "induction4" = "4", "induction5" = "5", "induction6" = "6", "induction7" = "7", "induction8" = "8")

ex1Tidy <- merge(ex1TidyT, ex1TidyD_short, by = c("SID", "item"))
ex1 <- ex1Tidy

#mark the positive valence and negative valence items to compare
ex1$valence<-ex1$item
ex1$valence<-recode_factor(ex1$valence, "1" = "positive", "2" = "negative", "3"  = "positive", "4" = "positive", "5" = "negative", "6" = "positive", "7" = "negative", "8" = "negative")
print(levels(ex1$valence))

#for simple effects in each age group
Child_5.6Tidy <- ex1 %>%
  filter(AgeGroup == "Child_5.6")
Child_7.8Tidy <- ex1 %>%
  filter(AgeGroup == "Child_7.8")
AdultTidy<-ex1 %>%
  filter(AgeGroup == "Adult")

#Tibbles for follow-up analyses
Representative <- ex1 %>%
  filter(Condition == "Representative")
Ideals <- ex1 %>%
  filter(Condition == "Best")
```

```{r, include = FALSE}
#Main analysis testing for effects of condition and age group on choice judgments in Study 1. This is a Linear Mixed Model, implemented with the lme4 package, using "lmer." This model tests for main and interactive effects of Condition and age group (between participants), and item Valence (within participants). Random instercepts are included for item and participant.

#choice
choice <- lmer(cbind(choice) ~ AgeGroup*Condition*valence + (1|SID) + (1|item), data = ex1)
choiceS<-summary(choice)
tableT<-as.data.frame(coef(choiceS))
tableT
ex1_ANOVAtableT<-as.data.frame(anova(choice, ddf = "Kenward-Roger", type = 3))
ex1_ANOVAtableT
```

```{r, include = FALSE}
#marginal means
meansAgeT<-effect("AgeGroup", choice)
meansAgeT<-as.data.frame(meansAgeT)
meansAgeT

meansCondT<-effect("Condition", choice)
meansCondT<-as.data.frame(meansCondT)
meansCondT

meansCondAgeT<-effect("AgeGroup*Condition", choice)
meansCondAgeT<-as.data.frame(meansCondAgeT)
meansCondAgeT

meansCondValT<-effect("Condition*valence", choice)
meansCondValT<-as.data.frame(meansCondValT)
meansCondValT

#cell means
meansCellT<-effect("AgeGroup*Condition*valence", choice)
meansCellT<-as.data.frame(meansCellT)
meansCellT

#simple main effects of age and valence in each condition

#Representative condition
Choice_Representative <- lmer(cbind(choice) ~ AgeGroup*valence + (1|SID) + (1|item), data = Representative)
Choice_RepresentativeS<-summary(Choice_Representative)
tableT_Representative<-as.data.frame(coef(Choice_RepresentativeS))
tableT_Representative
ex1_ANOVAtableT_Representative<-as.data.frame(anova(Choice_Representative, ddf = "Kenward-Roger", type = 3))
ex1_ANOVAtableT_Representative
```

```{r, include = FALSE}

#ideals condition
Choice_Ideals <- lmer(cbind(choice) ~ AgeGroup*valence + (1|SID) + (1|item), data = Ideals)
Choice_IdealsS<-summary(Choice_Ideals)
tableT_Ideals<-as.data.frame(coef(Choice_IdealsS))
tableT_Ideals
ex1_ANOVAtableT_Ideals<-as.data.frame(anova(Choice_Ideals, ddf = "Kenward-Roger", type = 3))
ex1_ANOVAtableT_Ideals
```

```{r, include = FALSE}
#marginal means
meansValT_Ideals<-effect("valence", Choice_Ideals)
meansValT_Ideals<-as.data.frame(meansValT_Ideals)
meansValT_Ideals

#Simple main effects within each age group
#5-6 year olds
choice5.6 <- lmer(cbind(choice) ~ Condition*valence + (1|SID) + (1|item), data = Child_5.6Tidy)
summary(choice5.6)
choiceS5.6<-summary(choice5.6)
tableT5.6<-as.data.frame(coef(choiceS5.6))
tableT5.6
ex1_ANOVAtableT5.6<-as.data.frame(anova(choice5.6, ddf = "Kenward-Roger", type = 3))
ex1_ANOVAtableT5.6
```

```{r, include = FALSE}
#marginal means
meansCondT5.6<-effect("Condition", choice5.6)
meansCondT5.6<-as.data.frame(meansCondT5.6)
meansCondT5.6

###7-8 year olds
choice7.8 <- lmer(cbind(choice) ~ Condition*valence + (1|SID) + (1|item), data = Child_7.8Tidy)
summary(choice7.8)
choiceS7.8<-summary(choice7.8)
tableT7.8<-as.data.frame(coef(choiceS7.8))
tableT7.8
ex1_ANOVAtableT7.8<-as.data.frame(anova(choice7.8, ddf = "Kenward-Roger", type = 3))
ex1_ANOVAtableT7.8

#marginal means
meansCondT7.8<-effect("Condition", choice7.8)
meansCondT7.8<-as.data.frame(meansCondT7.8)
meansCondT7.8

#cell means
meansCellT7.8<-effect("Condition*valence", choice7.8)
meansCellT7.8<-as.data.frame(meansCellT7.8)
meansCellT7.8

#Adults
choiceAdult <- lmer(cbind(choice) ~ Condition*valence + (1|SID) + (1|item), data = AdultTidy)
summary(choiceAdult)
choiceSAdult<-summary(choiceAdult)
tableTAdult<-as.data.frame(coef(choiceSAdult))
tableTAdult
ex1_ANOVAtableTAdult<-as.data.frame(anova(choiceAdult, ddf = "Kenward-Roger", type = 3))
ex1_ANOVAtableTAdult

#marginal means
meansValTAdult<-effect("valence", choiceAdult)
meansValTAdult<-as.data.frame(meansValTAdult)
meansValTAdult

meansCondTAdult<-effect("Condition", choiceAdult)
meansCondTAdult<-as.data.frame(meansCondTAdult)
meansCondTAdult

#cell means
meansCellTAdult<-effect("Condition*valence", choiceAdult)
meansCellTAdult<-as.data.frame(meansCellTAdult)
meansCellTAdult

# Induction - Study 1

#Main analysis testing for effects of condition and age group on induction judgments in Study 1. This is a Generalized Linear Mixed Model (GLMM), implemented with the lme4 package, using "glmer" to specify a binomial distribution, since it was forced-choice between average and extreme exemplars. This model tests for main and interactive effects of Condition and age group, and Valence (w/in participants). Random intercepts for item in participant are included. 

induction<-glmer(cbind(induction) ~ AgeGroup*Condition*valence + (1|SID) + (1|item), data = ex1, family=binomial)
inductionS<-summary(induction)
inductionS
tableD<-as.data.frame(coef(summary(induction)))
tableD
ex1_ANOVAtableD<-as.data.frame(car::Anova(induction, type = "III"))
ex1_ANOVAtableD

#marginal means
meansAgeD<-effect("AgeGroup", induction)
meansAgeD<-as.data.frame(meansAgeD)
meansAgeD
```

```{r, include = FALSE}
### Study 2 ###

#Read in the data file
#This file contains data from children and adults, in wide format
ex2Wide<- read_csv("Extremes2_Items.csv")

#Turn categorical variables into factors, and set reference categories and labels in a sensible and useful way

ex2Wide$Condition<-factor(ex2Wide$Condition, levels = c("1", "2"))
print(levels(ex2Wide$Condition))
ex2Wide$Condition<-recode_factor(ex2Wide$Condition, "1" = "Best", "2" = "Representative")

ex2Wide$Gender<-factor(ex2Wide$Gender, levels = c("1", "2"))
print(levels(ex2Wide$Gender))
ex2Wide$Gender<-recode_factor(ex2Wide$Gender, "1" = "Male", "2" = "Female", "3" = "Other")
print(levels(ex2Wide$Gender))

ex2Wide$AgeGroup<-factor(ex2Wide$AgeGroup, levels = c("1", "2", "3", "4"))
print(levels(ex2Wide$AgeGroup))
ex2Wide$AdultChild <- ex2Wide$AgeGroup
ex2Wide$AgeGroup<-recode_factor(ex2Wide$AgeGroup, "1" = "Child_5.6", "2" = "Child_7.8", "3" = "Child_9.10", "4" = "Adult")

ex2Wide$AdultChild <- ex2Wide$AgeGroup
ex2Wide$AdultChild<-recode_factor(ex2Wide$AdultChild, "Child_5.6" = "Child", "Child_7.8" = "Child", "Child_9.10" = "Child", "Adult" = "Adult")

ex2Wide_Child <- ex2Wide %>%
  filter(AdultChild == "Child")
ex2Wide_Adult <- ex2Wide %>%
  filter(AdultChild == "Adult")
ex2Wide_Child_5.6 <- ex2Wide_Child %>%
  filter(AgeGroup == "Child_5.6")
ex2Wide_Child_7.8 <- ex2Wide_Child %>%
  filter(AgeGroup == "Child_7.8")
ex2Wide_Child_9.10 <- ex2Wide_Child %>%
  filter(AgeGroup == "Child_9.10")

#Nature and science questionnaire info - make Community a factor to specify levels
ex2Wide$Community<-factor(ex2Wide$Community, levels = c("0", "1", "2"))
print(levels(ex2Wide$Community))
ex2Wide$Community<-recode_factor(ex2Wide$Community, "0" = "Urban", "1" = "Suburban", "2" = "Rural")
print(levels(ex2Wide$Community))

#Tidy the data
#For this dataset, that means making two separate tidy columns - one for choice (typicality or ideals) and one for induction

#Tidy - choice items
#we will add the discover items in a separate column after we tidy the choice into a column
ex2TidyT <- ex2Wide %>%
  gather(item, choice, choice1, choice2, choice3, choice4, choice5, choice6, choice7, choice8)

ex2TidyT$item<-factor(ex2TidyT$item, levels = c("choice1", "choice2", "choice3", "choice4", "choice5", "choice6", "choice7", "choice8"))
print(levels(ex2TidyT$item))
ex2TidyT$item<-recode_factor(ex2TidyT$item, "choice1" = "1", "choice2" = "2", "choice3" = "3", "choice4" = "4", "choice5" = "5", "choice6" = "6", "choice7" = "7", "choice8" = "8")

keepvarsT <- c("SID", "Condition", "ExactAge", "AgeGroup", "item", "choice")
ex2TidyT <- ex2TidyT[keepvarsT]

#Tidy - induction items
ex2TidyD <- ex2Wide %>%
  gather(item, induction, induction1, induction2, induction3, induction4, induction5, induction6, induction7, induction8)

keepvarsD <- c("SID", "item", "induction")
ex2TidyD_short <- ex2TidyD[keepvarsD]

ex2TidyD_short$item<-factor(ex2TidyD_short$item, levels = c("induction1", "induction2", "induction3", "induction4", "induction5", "induction6", "induction7", "induction8"))
print(levels(ex2TidyD_short$item))
ex2TidyD_short$item<-recode_factor(ex2TidyD_short$item, "induction1" = "1", "induction2" = "2", "induction3" = "3", "induction4" = "4", "induction5" = "5", "induction6" = "6", "induction7" = "7", "induction8" = "8")

ex2Tidy <- merge(ex2TidyT, ex2TidyD_short, by = c("SID", "item"))
ex2 <- ex2Tidy

#mark the positive valence and negative valence items to compare
ex2$valence<-ex2$item
ex2$valence<-recode_factor(ex2$valence, "1" = "positive", "2" = "negative", "3"  = "positive", "4" = "positive", "5" = "negative", "6" = "positive", "7" = "negative", "8" = "negative")
print(levels(ex2$valence))

#for simple effects in each age group
ex2_Child_5.6Tidy <- ex2 %>%
  filter(AgeGroup == "Child_5.6")
ex2_Child_7.8Tidy <- ex2 %>%
  filter(AgeGroup == "Child_7.8")
ex2_Child_9.10Tidy <- ex2 %>%
  filter(AgeGroup == "Child_9.10")
ex2_AdultTidy<-ex2 %>%
  filter(AgeGroup == "Adult")

#Tibbles for follow-up analyses
Representative2 <- ex2 %>%
  filter(Condition == "Representative")
Ideals2 <- ex2 %>%
  filter(Condition == "Best")

#Main analysis testing for effects of condition and age group on choice judgments in Study 2. This is a Linear Mixed Model, implemented with the lme4 package, using "lmer" to specify that item is repeated within participants. This model tests for main and interactive effects of Condition and age group. 

#choice
Choice2 <- lmer(cbind(choice) ~ AgeGroup*Condition*valence + (1|SID) + (1|item), data = ex2)
summary(Choice2)
choiceS_ex2<-summary(Choice2)
tableT_ex2<-as.data.frame(coef(choiceS_ex2))
tableT_ex2
ex2_ANOVAtableT<-as.data.frame(anova(Choice2, ddf = "Kenward-Roger", type = 3))
ex2_ANOVAtableT
```

```{r, include = FALSE}
#marginal means
meansAgeT_ex2<-effect("AgeGroup", Choice2)
meansAgeT_ex2<-as.data.frame(meansAgeT_ex2)
meansAgeT_ex2

meansCondT_ex2<-effect("Condition", Choice2)
meansCondT_ex2<-as.data.frame(meansCondT_ex2)
meansCondT_ex2

meansCondAgeT_ex2<-effect("AgeGroup*Condition", Choice2)
meansCondAgeT_ex2<-as.data.frame(meansCondAgeT_ex2)
meansCondAgeT_ex2

meansCondValT_ex2<-effect("Condition*valence", Choice2)
meansCondValT_ex2<-as.data.frame(meansCondValT_ex2)
meansCondValT_ex2

#cell means
meansCellT_ex2<-effect("AgeGroup*Condition*valence", Choice2)
meansCellT_ex2<-as.data.frame(meansCellT_ex2)
meansCellT_ex2

#simple main effects of age and valence in each condition

#Representative condition condition
Choice2_Representative <- lmer(cbind(choice) ~ AgeGroup*valence + (1|SID) + (1|item), data = Representative2)
summary(Choice2_Representative)
Choice2_RepresentativeS<-summary(Choice2_Representative)
tableT_Representative<-as.data.frame(coef(Choice2_RepresentativeS))
tableT_Representative
ex2_ANOVAtableT_Representative<-as.data.frame(anova(Choice2_Representative, ddf = "Kenward-Roger", type = 3))
ex2_ANOVAtableT_Representative

#Ideal condition
Choice2_Ideals <- lmer(cbind(choice) ~ AgeGroup*valence + (1|SID) + (1|item), data = Ideals2)
summary(Choice2_Ideals)
Choice2_IdealsS<-summary(Choice2_Ideals)
tableT_Ideals<-as.data.frame(coef(Choice2_IdealsS))
tableT_Ideals
ex2_ANOVAtableT_Ideals<-as.data.frame(anova(Choice2_Ideals, ddf = "Kenward-Roger", type = 3))
ex2_ANOVAtableT_Ideals

#marginal means
meansValT_Ideals<-effect("valence", Choice2_Ideals)
meansValT_Ideals<-as.data.frame(meansValT_Ideals)
meansValT_Ideals

#Simple main effects within each age group
#5-6 year olds
choice2_5.6 <- lmer(cbind(choice) ~ Condition*valence + (1|SID) + (1|item), data = ex2_Child_5.6Tidy)
summary(choice2_5.6)
choiceS5.6<-summary(choice2_5.6)
tableT5.6<-as.data.frame(coef(choiceS5.6))
tableT5.6
ex2_ANOVAtableT5.6<-as.data.frame(anova(choice2_5.6, ddf = "Kenward-Roger", type = 3))
ex2_ANOVAtableT5.6
```

```{r, include = FALSE}
#marginal means
meansCondT_ex25.6<-effect("Condition", choice2_5.6)
meansCondT_ex25.6<-as.data.frame(meansCondT_ex25.6)
meansCondT_ex25.6

#cell means
meansCellT_ex25.6<-effect("Condition*valence", choice2_5.6)
meansCellT_ex25.6<-as.data.frame(meansCellT_ex25.6)
meansCellT_ex25.6

###7-8 year olds
choice2_7.8 <- lmer(cbind(choice) ~ Condition*valence + (1|SID) + (1|item), data = ex2_Child_7.8Tidy)
summary(choice2_7.8)
choiceS7.8<-summary(choice2_7.8)
tableT7.8<-as.data.frame(coef(choiceS7.8))
tableT7.8
ex2_ANOVAtableT7.8<-as.data.frame(anova(choice2_7.8, ddf = "Kenward-Roger", type = 3))
ex2_ANOVAtableT7.8

#marginal means
meansCondT_ex27.8<-effect("Condition", choice2_7.8)
meansCondT_ex27.8<-as.data.frame(meansCondT_ex27.8)
meansCondT_ex27.8

#cell means
meansCellT_ex27.8<-effect("Condition*valence", choice2_7.8)
meansCellT_ex27.8<-as.data.frame(meansCellT_ex27.8)
meansCellT_ex27.8

###9-10 year olds
choice2_9.10 <- lmer(cbind(choice) ~ Condition*valence + (1|SID) + (1|item), data = ex2_Child_9.10Tidy)
summary(choice2_9.10)
choiceS9.10<-summary(choice2_9.10)
tableT9.10<-as.data.frame(coef(choiceS9.10))
tableT9.10
ex2_ANOVAtableT9.10<-as.data.frame(anova(choice2_9.10, ddf = "Kenward-Roger", type = 3))
ex2_ANOVAtableT9.10

#marginal means
meansCondT_ex29.10<-effect("Condition", choice2_9.10)
meansCondT_ex29.10<-as.data.frame(meansCondT_ex29.10)
meansCondT_ex29.10

#cell means
meansCellT_ex29.10<-effect("Condition*valence", choice2_9.10)
meansCellT_ex29.10<-as.data.frame(meansCellT_ex29.10)
meansCellT_ex29.10

#Adults
choiceAdult_ex2 <- lmer(cbind(choice) ~ Condition*valence + (1|SID) + (1|item), data = ex2_AdultTidy)
summary(choiceAdult_ex2)
choiceSAdult_ex2<-summary(choiceAdult_ex2)
tableTAdult_ex2<-as.data.frame(coef(choiceSAdult_ex2))
tableTAdult_ex2
ex2_ANOVAtableTAdult<-as.data.frame(anova(choiceAdult_ex2, ddf = "Kenward-Roger", type = 3))
ex2_ANOVAtableTAdult

#marginal means
meansCondT_ex2Adult<-effect("Condition", choiceAdult)
meansCondT_ex2Adult<-as.data.frame(meansCondT_ex2Adult)
meansCondT_ex2Adult

#cell means
meansCellT_ex2Adult<-effect("Condition*valence", choiceAdult)
meansCellT_ex2Adult<-as.data.frame(meansCellT_ex2Adult)
meansCellT_ex2Adult

# Induction - Study 2
induction2<-glmer(cbind(induction) ~ AgeGroup*Condition*valence + (1|SID) + (1|item), data = ex2, family=binomial)
induction2S<-summary(induction2)
induction2S
tableD<-as.data.frame(coef(summary(induction2)))
tableD
ex2_ANOVAtableD<-as.data.frame(car::Anova(induction2, type = "III"))
ex2_ANOVAtableD

#marginal means
meansAgeD_ex2<-effect("AgeGroup", induction2)
meansAgeD_ex2<-as.data.frame(meansAgeD_ex2)
meansAgeD_ex2
```

```{r, echo = FALSE, warning = FALSE}

#Create Figure 2 - choice, studies 1 and 2
#Study 1
Study_1<-ggplot() +
  geom_hline(mapping = NULL, data = NULL, colour = "#F9610A", yintercept = 3, na.rm = FALSE,
  show.legend = NA) +
  geom_point(data = meansCondAgeT, aes(x = AgeGroup, y = fit, colour = Condition, shape = Condition), stat = "identity", size = 4) +
    scale_x_discrete(limits=c("Child_5.6", "Child_7.8", "Adult"), labels = c("Child_5.6" = "5-6", "Child_7.8" = "7-8", "Adult" = "Adults")) +
  ylim(1,5.2) +
  geom_jitter(data = ex1, aes(x = AgeGroup, y = choice, colour = Condition), width = .30, height = .00, alpha = .5, shape = 124, size = 1) +
  ggtitle("Study 1") +
  theme_few() +
  theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10)) +
  theme(legend.position="bottom", legend.direction="horizontal") +
  theme(legend.title = element_blank()) +
  scale_shape_manual(values=c(17, 19)) +
  scale_color_manual(values=c("#4C9900", "#0066CC"), guide = guide_legend(reverse = FALSE)) +
labs(x="age group", y="Average exemplar selected")

#Study 2
Study_2<-ggplot() +
  geom_hline(mapping = NULL, data = NULL, colour = "#F9610A", yintercept = 3, na.rm = FALSE,
  show.legend = NA) +
  geom_point(data = meansCondAgeT_ex2, aes(x = AgeGroup, y = fit, colour = Condition, shape = Condition), stat = "identity", size = 4) +
    scale_x_discrete(limits=c("Child_5.6", "Child_7.8", "Child_9.10", "Adult"), labels = c("Child_5.6" = "5-6", "Child_7.8" = "7-8", "Child_9.10" = "9-10", "Adult" = "Adults")) +
  ylim(1,5.2) +
  geom_jitter(data = ex2, aes(x = AgeGroup, y = choice, colour = Condition), width = .30, height = .00, alpha = .5, shape = 124, size = 1) +
  ggtitle("Study 2") +
  theme_few() +
  theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10)) +
  theme(legend.position="bottom", legend.direction="horizontal") +
  theme(legend.title = element_blank()) +
  scale_shape_manual(values=c(17, 19)) +
  scale_color_manual(values=c("#4C9900", "#0066CC"), guide = guide_legend(reverse = FALSE)) +
  labs(x="age group", y=NULL)

#Comine into one figure
require(cowplot)
Figure2 <- plot_grid(Study_1, Study_2)
save_plot("exfig2.png", Figure2, base_aspect_ratio = 1.3)
```

```{r, include = FALSE}
### Study 3 ###

#Read in the data file
#This file contains data from children and adults, in wide format
ex3Wide<- read_csv("Extremes3_Items.csv")

#Turn categorical variables into factors, and set reference categories and labels in a sensible and useful way

ex3Wide$Contrast<-factor(ex3Wide$Contrast, levels = c("0", "1"))
print(levels(ex3Wide$Contrast))
ex3Wide$Contrast<-recode_factor(ex3Wide$Contrast, "0" = "No Contrast", "1" = "Contrast")

ex3Wide$Functional<-factor(ex3Wide$Functional, levels = c("0", "1"))
print(levels(ex3Wide$Functional))
ex3Wide$Functional<-recode_factor(ex3Wide$Functional, "0" = "Nonfunctional", "1" = "Functional")

ex3Wide$Gender<-factor(ex3Wide$Gender, levels = c("1", "2"))
print(levels(ex3Wide$Gender))
ex3Wide$Gender<-recode_factor(ex3Wide$Gender, "1" = "Male", "2" = "Female", "3" = "Other")

ex3Wide$AgeGroup<-factor(ex3Wide$AgeGroup, levels = c("1", "2", "3"))
print(levels(ex3Wide$AgeGroup))
ex3Wide$AdultChild <- ex3Wide$AgeGroup

ex3Wide$AgeGroup<-recode_factor(ex3Wide$AgeGroup, "1" = "Child_5.6", "2" = "Child_7.8", "3" = "Adult")

ex3Wide$AdultChild<-recode_factor(ex3Wide$AdultChild, "1" = "Child", "2" = "Child", "3" = "Adult")

#Make Community a factor to specify levels
ex3Wide$Community<-factor(ex3Wide$Community, levels = c("0", "1", "2"))
print(levels(ex3Wide$Community))
ex3Wide$Community<-recode_factor(ex3Wide$Community, "0" = "Urban", "1" = "Suburban", "2" = "Rural")
print(levels(ex3Wide$Community))

#Now let's make some subsets of this dataset for participants info
ex3Wide_Child_5.6 <- ex3Wide %>%
  filter(AgeGroup == "Child_5.6")
ex3Wide_Child_7.8 <- ex3Wide %>%
  filter(AgeGroup == "Child_7.8")
ex3Wide_Adult<-ex3Wide %>%
  filter(AgeGroup == "Adult")
ex3Wide_Child <- ex3Wide %>%
  filter(AdultChild == "Child")
```

```{r, include = FALSE}
#Now we need to tidy the data
#For this dataset, that means making two separate tidy columns - one for representative and one for discover (induction)

#Tidy - representative (typicality) items
#we will add the discover items in a separate column after we tidy the representative into a column
ex3TidyT <- ex3Wide %>%
  gather(item, representative, typ1, typ2, typ3, typ4, typ5, typ6, typ7, typ8)

ex3TidyT$item<-factor(ex3TidyT$item, levels = c("typ1", "typ2", "typ3", "typ4", "typ5", "typ6", "typ7", "typ8"))
print(levels(ex3TidyT$item))
ex3TidyT$item<-recode_factor(ex3TidyT$item, "typ1" = "1", "typ2" = "2", "typ3" = "3", "typ4" = "4", "typ5" = "5", "typ6" = "6", "typ7" = "7", "typ8" = "8")

keepvarsT <- c("SID", "Contrast", "Functional", "ExactAge", "AgeGroup", "Gender", "item", "representative")
ex3TidyT <- ex3TidyT[keepvarsT]

#Tidy - discover items
ex3TidyD <- ex3Wide %>%
  gather(item, induction, discover1, discover2, discover3, discover4, discover5, discover6, discover7, discover8)

keepvarsD <- c("SID", "item", "induction")
ex3TidyD_short <- ex3TidyD[keepvarsD]

ex3TidyD_short$item<-factor(ex3TidyD_short$item, levels = c("discover1", "discover2", "discover3", "discover4", "discover5", "discover6", "discover7", "discover8"))
print(levels(ex3TidyD_short$item))
ex3TidyD_short$item<-recode_factor(ex3TidyD_short$item, "discover1" = "1", "discover2" = "2", "discover3" = "3", "discover4" = "4", "discover5" = "5", "discover6" = "6", "discover7" = "7", "discover8" = "8")

ex3Tidy <- merge(ex3TidyT, ex3TidyD_short, by = c("SID", "item"))
ex3 <- ex3Tidy

#Copy the iduction and typicality items into raw responses (for items analysis) before reverse-coding
ex3$representativeRaw<-ex3$representative
ex3$inductionRaw<-ex3$induction

#Mark which items get reverse-coded with a 1 (others with a 0)
ex3$reverse<-ex3$item
ex3$reverse<-recode_factor(ex3$reverse, "1" = "1", "2" = "0", "3" = "0", "4" = "1", "5" = "1", "6" = "0", "7" = "0", "8" = "1")
print(levels(ex3$reverse))

#Reverse-code representative and induction responses for the 4 items presented in reversed direction, to allow comparison across items
ex3$representative_r<-recode(ex3$representative, "5=1; 4=2; 3=3; 2=4; 1=5")
ex3$induction_r<-recode(ex3$induction, "5=1; 4=2; 3=3; 2=4; 1=5")
ex3$representative[ex3$reverse == "1"] <- ex3$representative_r[ex3$reverse == "1"]
ex3$induction[ex3$reverse == "1"] <- ex3$induction_r[ex3$reverse == "1"]

#Mark the direction of progression in reverse-coded responses (most-least = increase in amount of varied property progresses from left to right, least-most = progresses from right to left). Since reverse-coding made all scales organized so the the most functionally ideal is the highest number (i.e., progressing to the right), so scales with a "most-least" direction are those in which the functional ideal has least rather than most of the varied property.

ex3$direction<-ex3$item
ex3$direction<-recode_factor(ex3$item, "1" = "least-most", "2" = "least-most", "3"  = "least-most", "4" = "most-least", "5" = "most-least", "6" = "most-least", "7" = "least-most", "8" = "most-least")
print(levels(ex3$direction))

#For simple effects in each age group
ex3_Child_5.6Tidy <- ex3 %>%
  filter(AgeGroup == "Child_5.6")
ex3_Child_7.8Tidy <- ex3 %>%
  filter(AgeGroup == "Child_7.8")
ex3_AdultTidy<-ex3 %>%
  filter(AgeGroup == "Adult")

#Tibbles for follow-up analyses
Functional <- ex3 %>%
  filter(Functional == "Functional")
NoFunctional <- ex3 %>%
  filter(Functional == "Nonfunctional")
Contrast <- ex3 %>%
  filter(Contrast == "Contrast")
NoContrast <- ex3 %>%
  filter(Contrast == "No Contrast")
```

```{r, include = FALSE}
#Main analysis testing for effects of age, functional information, contrast category, and scale direction in Study 3. This is a Linear Mixed Model, implemented with the lme4 package. Random intercepts are included for item and participant.

#Representativeness (representative)
Representative <- lmer(cbind(representative) ~ AgeGroup*Contrast*Functional + (1|SID) + (1|item), data = ex3)
RepresentativeS<-summary(Representative)
tableT<-as.data.frame(coef(RepresentativeS))
tableT
ex3_ANOVAtableT<-as.data.frame(anova(Representative, ddf = "Kenward-Roger", type = 3))
ex3_ANOVAtableT
```

```{r, include = FALSE}
#marginal means
meansAgeT_ex3<-effect("AgeGroup", Representative)
meansAgeT_ex3<-as.data.frame(meansAgeT_ex3)
meansAgeT_ex3

meansContrastT<-effect("Contrast", Representative)
meansContrastT<-as.data.frame(meansContrastT)
meansContrastT

meansFuncT<-effect("Functional", Representative)
meansFuncT<-as.data.frame(meansFuncT)
meansFuncT

meansFunAgeT<-effect("AgeGroup*Functional", Representative)
meansFunAgeT<-as.data.frame(meansFunAgeT)
meansFunAgeT

#cell means
meansCellT_ex3<-effect("AgeGroup*Contrast*Functional", Representative)
meansCellT_ex3<-as.data.frame(meansCellT_ex3)
meansCellT_ex3

#Simple main effects within each age group
#5-6 year olds
Representative5.6 <- lmer(cbind(representative) ~ Contrast*Functional + (1|SID) + (1|item), data = ex3_Child_5.6Tidy)
RepresentativeS5.6<-summary(Representative5.6)
tableT5.6<-as.data.frame(coef(RepresentativeS5.6))
tableT5.6
ex3_ANOVAtableT5.6<-as.data.frame(anova(Representative5.6, ddf = "Kenward-Roger", type = 3))
ex3_ANOVAtableT5.6
```

```{r, include = FALSE}
###7-8 year olds
Representative7.8 <- lmer(cbind(representative) ~ Contrast*Functional + (1|SID) + (1|item), data = ex3_Child_7.8Tidy)
summary(Representative7.8)
RepresentativeS7.8<-summary(Representative7.8)
tableT7.8<-as.data.frame(coef(RepresentativeS7.8))
tableT7.8
ex3_ANOVAtableT7.8<-as.data.frame(anova(Representative7.8, ddf = "Kenward-Roger", type = 3))
ex3_ANOVAtableT7.8

#Adults
RepresentativeAdult <- lmer(cbind(representative) ~ Contrast*Functional + (1|SID) + (1|item), data = ex3_AdultTidy)
summary(RepresentativeAdult)
RepresentativeSAdult<-summary(RepresentativeAdult)
tableTAdult<-as.data.frame(coef(RepresentativeSAdult))
tableTAdult
ex3_ANOVAtableTAdult<-as.data.frame(anova(RepresentativeAdult, ddf = "Kenward-Roger", type = 3))
ex3_ANOVAtableTAdult
```

```{r, include = FALSE}
#Discover items (sample selection: "let's discover new things about [animal]")
Discover <- lmer(cbind(induction) ~ AgeGroup*Contrast*Functional + (1|SID) + (1|item), data = ex3)
DiscoverS<-summary(Discover)
tableD<-as.data.frame(coef(DiscoverS))
tableD
ex3_ANOVAtableD<-as.data.frame(anova(Discover, ddf = "Kenward-Roger", type = 3))
ex3_ANOVAtableD
```

```{r, include = FALSE}
#marginal means
meansAgeD_ex3<-effect("AgeGroup", Discover)
meansAgeD_ex3<-as.data.frame(meansAgeD_ex3)
meansAgeD_ex3

meansContrastD<-effect("Contrast", Discover)
meansContrastD<-as.data.frame(meansContrastD)
meansContrastD

meansFuncD<-effect("Functional", Discover)
meansFuncD<-as.data.frame(meansFuncD)
meansFuncD

meansCondD<-effect("Contrast*Functional", Discover)
meansCondD<-as.data.frame(meansCondD)
meansCondD

meansFunAgeD<-effect("AgeGroup*Functional", Discover)
meansFunAgeD<-as.data.frame(meansFunAgeD)
meansFunAgeD

#cell means
meansCellD_ex3<-effect("AgeGroup*Contrast*Functional", Discover)
meansCellD_ex3<-as.data.frame(meansCellD_ex3)
meansCellD_ex3

#Simple main effects within each age group
#5-6 year olds
Discover5.6 <- lmer(cbind(induction) ~ Contrast*Functional + (1|SID) + (1|item), data = ex3_Child_5.6Tidy)
DiscoverS5.6<-summary(Discover5.6)
tableD5.6<-as.data.frame(coef(DiscoverS5.6))
tableD5.6
ex3_ANOVAtableD5.6<-as.data.frame(anova(Discover5.6, ddf = "Kenward-Roger", type = 3))
ex3_ANOVAtableD5.6
```

```{r, include = FALSE}
#7-8 year olds
Discover7.8 <- lmer(cbind(induction) ~ Contrast*Functional*direction + (1|SID) + (1|item), data = ex3_Child_7.8Tidy)
summary(Discover7.8)
DiscoverS7.8<-summary(Discover7.8)
tableD7.8<-as.data.frame(coef(DiscoverS7.8))
tableD7.8
ex3_ANOVAtableD7.8<-as.data.frame(anova(Discover7.8, ddf = "Kenward-Roger", type = 3))
ex3_ANOVAtableD7.8

#Adults
DiscoverAdult <- lmer(cbind(induction) ~ Contrast*Functional*direction + (1|SID) + (1|item), data = ex3_AdultTidy)
summary(DiscoverAdult)
DiscoverSAdult<-summary(DiscoverAdult)
tableDAdult<-as.data.frame(coef(DiscoverSAdult))
tableDAdult
ex3_ANOVAtableDAdult<-as.data.frame(anova(DiscoverAdult, ddf = "Kenward-Roger", type = 3))
ex3_ANOVAtableDAdult
```

```{r, echo = FALSE, warning = FALSE}
#Figure for Study 3, with representativeness and discover (sample selection) choices
#Create Figure 4a--representativeness, collapsed across contrast since this was not significant in a meaningful way for the current hypotheses
Figure4a<-ggplot() +
  geom_hline(mapping = NULL, data = NULL, colour ="#F9610A", yintercept = 3, na.rm = FALSE,
  show.legend = NA) +
  geom_point(data = meansFunAgeT, aes(x = AgeGroup, y = fit, colour = Functional, shape = Functional), stat = "identity", size = 4) +
    scale_x_discrete(limits=c("Child_5.6", "Child_7.8", "Adult"), labels = c("Child_5.6" = "5-6 yrs","Child_7.8" = "7-8 yrs", "Adult" = "Adults")) +
  ylim(1,5) +
  geom_jitter(data = ex3, aes(x = AgeGroup, y = representative, colour = Functional), width = .30, height = .00, alpha = .5, shape = 124, size = 1) +
  ggtitle("Representative") +
  theme_few() +
  theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10)) +
  theme(legend.position="bottom", legend.direction="horizontal") +
  theme(legend.title = element_blank()) +
  scale_shape_manual(values=c(17, 19)) +
  scale_color_manual(values=c("#4C9900", "#0066CC"), guide = guide_legend(reverse = FALSE)) +
  labs(x="age group", y="Average exemplar selected")

#Create Figure 4b--discover (sample selection), also collapsed across contrast
Figure4b<-ggplot() +
  geom_hline(mapping = NULL, data = NULL, colour = "#F9610A", yintercept = 3, na.rm = FALSE,
  show.legend = NA) +
  geom_point(data = meansFunAgeD, aes(x = AgeGroup, y = fit, colour = Functional, shape = Functional), stat = "identity", size = 4) +
  scale_x_discrete(limits=c("Child_5.6", "Child_7.8", "Adult"), labels = c("Child_5.6" = "5-6","Child_7.8" = "7-8", "Adult" = "Adults")) +
  ylim(1,5) +
  geom_jitter(data = ex3, aes(x = AgeGroup, y = induction, colour = Functional), width = .30, height = .00, alpha = .5, shape = 124, size = 1) +
  ggtitle("Sample Selection") +
  theme_few() +
  theme(axis.text=element_text(size=10),
        axis.title=element_text(size=10)) +
  theme(legend.position="bottom", legend.direction="horizontal") +
  theme(legend.title = element_blank()) +
  scale_shape_manual(values=c(17, 19)) +
  scale_color_manual(values=c("#4C9900", "#0066CC"), guide = guide_legend(reverse = FALSE)) +
    labs(x="age group", y=NULL)

#Combine plots into one figure
require(cowplot)
Figure4 <- plot_grid(Figure4a, Figure4b, align = "h")
save_plot("exfig4.png", Figure4, base_aspect_ratio = 1.3)
```

```{r, include = FALSE}
#Make a common dataset with demographics info for Studies 1-3 in order to describe the population as briefly as possible
#First, select only the variables we need (and that are common to all three)
ex1Demo <- ex1Wide
keepvarsDemo <- c("SID", "ExactAge", "AdultChild", "RE_code", "Hispanic", "Community", "State")
ex1Demo <- ex1Demo[keepvarsDemo]

ex2Demo <- ex2Wide
ex2Demo <- ex2Demo[keepvarsDemo]

ex3Demo <- ex3Wide
ex3Demo <- ex3Demo[keepvarsDemo]

#Combine into one dataset
Demo <- rbind(ex1Demo, ex2Demo, ex3Demo)

#Make subsets for adults and children
Demo_Child <- Demo %>%
  filter(AdultChild == "Child")
Demo_Adult <- Demo %>%
  filter(AdultChild == "Adult")
```
**Study 1**
**Participants.** In Study 1, children were ages 5-6 (*N* = `r nrow(ex1Wide_Child_5.6)`; *M* age = `r round(mean(ex1Wide_Child_5.6$ExactAge, na.rm = TRUE), 2)`; `r sum(ex1Wide_Child_5.6$Gender == "Male")` male) and 7-8 (*N* = `r nrow(ex1Wide_Child_7.8)`; *M* age = `r round(mean(ex1Wide_Child_7.8$ExactAge, na.rm = TRUE), 2)`; `r sum(ex1Wide_Child_7.8$Gender == "Male")` male). 

Across studies, children were `r round((sum(Demo_Child$RE_code == "White", na.rm = TRUE))/nrow(Demo_Child)*100, 0)`% White, `r round((sum(Demo_Child$RE_code == "Black", na.rm = TRUE))/nrow(Demo_Child)*100, 0)`% Black, `r round((sum(Demo_Child$RE_code == "Asian", na.rm = TRUE))/nrow(Demo_Child)*100, 0)`% Asian, `r round((sum(Demo_Child$RE_code == "MidEast", na.rm = TRUE))/nrow(Demo_Child)*100, 0)`% Middle Eastern or North African, and `r round((sum(Demo_Child$RE_code == "Mix", na.rm = TRUE))/nrow(Demo_Child)*100, 0)`% more than one race; `r round((sum(Demo_Child$Hispanic == "1", na.rm = TRUE))/nrow(Demo_Child)*100, 0)`% were Hispanic (of any race); `r round((sum(Demo_Child$RE_code == "Unreported", na.rm = TRUE))/nrow(Demo_Child)*100, 0)`% of parents did not report demographics information. Children gave oral assent and parents provided written consent.

Adults in Study 1 (*N* = `r nrow(ex1Wide_Adult)`, *M* age = `r round(mean(ex1Wide_Adult$ExactAge, na.rm = TRUE), 2)`; `r sum(ex1Wide_Adult$Gender == "Male")` male) were recruited with Amazon Mechanical Turk and tested using Qualtrics. Across studies, adults were `r round((sum(Demo_Adult$RE_code == "White"))/nrow(Demo_Adult)*100, 0)`% White, `r round((sum(Demo_Adult$RE_code == "Black"))/nrow(Demo_Adult)*100, 0)`% Black, `r round((sum(Demo_Adult$RE_code == "Asian"))/nrow(Demo_Adult)*100, 0)`% Asian, `r round((sum(Demo_Adult$RE_code == "NativeAm"))/nrow(Demo_Adult)*100, 0)`% Native American or Alaskan Native, and `r round((sum(Demo_Adult$RE_code == "Mix"))/nrow(Demo_Adult)*100, 0)`% more than one race; `r round((sum(Demo_Adult$Hispanic == "1", na.rm = TRUE))/nrow(Demo_Adult)*100, 0)`% were Hispanic (of any race). Adults provided written consent. All study procedures were approved by the Institutional Review Board of New York University. 

**Results**
Participants of all ages selected extreme exemplars as the “best” of their categories (e.g., they gave a prize to the fastest cheetah). Judgments of which exemplars were most representative changed across age, however. Younger children selected extreme exemplars as most representative, and these selections declined with age in favor of more average exemplars. The age x condition interaction was reliable (*F*(`r ex1_ANOVAtableT[4,3]`, `r ex1_ANOVAtableT[4,4]`) = `r round((ex1_ANOVAtableT[4,5]),2)`, *p* `r format.pval((ex1_ANOVAtableT[4,6]), digits = 2, eps = .001)`) as were subsumed main effects of age (*F*(`r ex1_ANOVAtableT[1,3]`, `r ex1_ANOVAtableT[1,4]`) = `r round((ex1_ANOVAtableT[1,5]),2)`, *p* `r format.pval((ex1_ANOVAtableT[1,6]), digits = 2, eps = .001)`) and condition (*F*(`r ex1_ANOVAtableT[2,3]`, `r ex1_ANOVAtableT[2,4]`) = `r round((ex1_ANOVAtableT[2,5]),2)`, *p* `r format.pval((ex1_ANOVAtableT[2,6]), digits = 2, eps = .001)`). Younger children were equally likely to pick extreme exemplars when asked to select the ideal as when asked to pick the most representative (*F*(`r ex1_ANOVAtableT5.6[1,3]`, `r ex1_ANOVAtableT5.6[1,4]`) = `r round((ex1_ANOVAtableT5.6[1,5]),2)`, *p* = `r format.pval((ex1_ANOVAtableT5.6[1,6]), digits = 2, eps = .001)`), whereas children ages 7-8 (*F*(`r ex1_ANOVAtableT7.8[1,3]`, `r ex1_ANOVAtableT7.8[1,4]`) = `r round((ex1_ANOVAtableT7.8[1,5]),2)`, *p* `r format.pval((ex1_ANOVAtableT7.8[1,6]), digits = 2, eps = .001)`) and adults (*F*(`r ex1_ANOVAtableTAdult[1,3]`, `r ex1_ANOVAtableTAdult[1,4]`) = `r round((ex1_ANOVAtableTAdult[1,5]),2)`, *p* `r format.pval((ex1_ANOVAtableTAdult[1,6]), digits = 2, eps = .001)`) selected more extreme exemplars when asked to select the ideal than when asked to pick the most representative.

Younger children also viewed extreme exemplars as providing more generalizable information. They were more likely to say, for example, that an unseen cheetah would be like the fastest cheetah (coded as 1) than the average-speed cheetah (coded as 0; *M* = `r round((meansAgeD[1,2]),2)`, 95% CI [`r round((meansAgeD[1,4]),2)`, `r round((meansAgeD[1,5]),2)`]. This tendency declined with age: Older children and adults more often generalized from average (instead of extreme) exemplars (ages 7-8: *M* = `r round((meansAgeD[2,2]),2)`, 95% CI [`r round((meansAgeD[2,4]),2)`, `r round((meansAgeD[2,5]),2)`]; adults: *M* = `r round((meansAgeD[3,2]),2)`, 95% CI [`r round((meansAgeD[3,4]),2)`, `r round((meansAgeD[3,5]),2)`]), Wald $X^2$(`r ex1_ANOVAtableD[2,2]`) = `r round((ex1_ANOVAtableD[2,1]),2)`, *p* `r format.pval((ex1_ANOVAtableD[2,3]), digits = 2, eps = .001)`.

Considering the valence of the properties (e.g., whether an extreme value was desirable or not from a human perspective), participants were slightly less likely to choose extreme values as ideal when the properties were objectionable from a human perspective (*F*(`r ex1_ANOVAtableT_Ideals[2,3]`, `r ex1_ANOVAtableT_Ideals[2,4]`) = `r round((ex1_ANOVAtableT_Ideals[2,5]),2)`, *p* = `r format.pval((ex1_ANOVAtableT_Ideals[2,6]), digits = 2, eps = .001)`; negative, e.g., the stinkiest skunk, *M* = `r round((meansValT_Ideals[2,2]),2)`, 95% CI [`r round((meansValT_Ideals[2,4]),2)`, `r round((meansValT_Ideals[2,5]),2)`]; positive, e.g., the fastest cheetah, *M* = `r round((meansValT_Ideals[1,2]),2)`, 95% CI [`r round((meansValT_Ideals[1,4]),2)`, `r round((meansValT_Ideals[1,5]),2)`]). The effect of valence on judgments about ideals did not interact with age, *F*(`r ex1_ANOVAtableT_Ideals[3,3]`, `r ex1_ANOVAtableT_Ideals[3,4]`) = `r round((ex1_ANOVAtableT_Ideals[3,5]),2)`, *p* = `r format.pval((ex1_ANOVAtableT_Ideals[3,6]), digits = 2, eps = .001)`). Critically, there were no effects of valence on representativeness choices (*F*(`r ex1_ANOVAtableT_Representative[2,3]`, `r ex1_ANOVAtableT_Representative[2,4]`) = `r round((ex1_ANOVAtableT_Representative[2,5]),2)`, *p* = `r format.pval((ex1_ANOVAtableT_Representative[2,6]), digits = 2, eps = .001)`) or induction (Wald $X^2$(`r ex1_ANOVAtableD[4,2]`) = `r round((ex1_ANOVAtableD[4,1]),2)`, *p* = `r format.pval((ex1_ANOVAtableD[4,3]), digits = 2, eps = .001)`).

Overall, Study 1 revealed that younger children view exemplars with more extreme values of characteristic properties as representative and informative. Older children and adults, in contrast, view more average exemplars in this manner. 

```{r}
Figure2
```
**Figure 2.** Average exemplar selected (as the best or most representative) in Studies 1-2 (N = 363), by age group and condition. Higher values represent exemplars that are more extreme on a characteristic property (e.g., faster cheetahs, stinkier skunks). Exemplar 3 represents the average, marked by the orange line. Large shapes are means by age group and condition; small lines are responses to individual trials.

**Study 2**
**Participants**
In Study 2, children were ages 5-6 (*N* = `r nrow(ex2Wide_Child_5.6)`; *M* age = `r round(mean(ex2Wide_Child_5.6$ExactAge, na.rm = TRUE), 2)`; `r sum(ex2Wide_Child_5.6$Gender == "Male")` male) 7-8 (*N* = `r nrow(ex2Wide_Child_7.8)`; *M* age = `r round(mean(ex2Wide_Child_7.8$ExactAge, na.rm = TRUE), 2)`; `r sum(ex2Wide_Child_7.8$Gender == "Male")` male) and 9-10 (*N* = `r nrow(ex2Wide_Child_9.10)`; *M* age = `r round(mean(ex2Wide_Child_9.10$ExactAge, na.rm = TRUE), 2)`; `r sum(ex2Wide_Child_9.10$Gender == "Male")` male). Adults in Study 2 (`r nrow(ex2Wide_Adult)`, *M* age = `r round(mean(ex2Wide_Adult$ExactAge, na.rm = TRUE), 2)`; `r sum(ex2Wide_Adult$Gender == "Male", na.rm = TRUE)` male) were recruited using Amazon Mechanical Turk and tested using Qualtrics.

**Results**
The pattern of results was identical to Study 1 (Figure 2). Participants of all ages selected extreme exemplars as “the best” of their categories, but judgments of which exemplars were most representative changed across age: Younger children selected extreme exemplars as most representative, and such selections declined with age in favor of more average exemplars. The age x condition interaction was reliable (*F*(`r ex2_ANOVAtableT[4,3]`, `r ex2_ANOVAtableT[4,4]`) = `r round((ex2_ANOVAtableT[4,5]),2)`, *p* `r format.pval((ex2_ANOVAtableT[4,6]), digits = 2, eps = .001)`) as were subsumed main effects of age (*F*(`r ex2_ANOVAtableT[1,3]`, `r ex2_ANOVAtableT[1,4]`) = `r round((ex2_ANOVAtableT[1,5]),2)`, *p* `r format.pval((ex2_ANOVAtableT[1,6]), digits = 2, eps = .001)`) and condition (*F*(`r ex2_ANOVAtableT[2,3]`, `r ex2_ANOVAtableT[2,4]`) = `r round((ex2_ANOVAtableT[2,5]),2)`, *p* `r format.pval((ex2_ANOVAtableT[2,6]), digits = 2, eps = .001)`). Younger children were just as likely to pick extreme exemplars when asked to select the ideal as when asked to pick the most representative (*F*(`r ex2_ANOVAtableT5.6[1,3]`, `r ex2_ANOVAtableT5.6[1,4]`) = `r round((ex2_ANOVAtableT5.6[1,5]),2)`, *p* = `r format.pval((ex2_ANOVAtableT5.6[1,6]), digits = 2, eps = .001)`), whereas children ages 7-8 (*F*(`r ex2_ANOVAtableT7.8[1,3]`, `r ex2_ANOVAtableT7.8[1,4]`) = `r round((ex2_ANOVAtableT7.8[1,5]),2)`, *p* = `r format.pval((ex2_ANOVAtableT7.8[1,6]), digits = 2, eps = .001)`), ages 9-10 (*F*(`r ex2_ANOVAtableT9.10[1,3]`, `r ex2_ANOVAtableT9.10[1,4]`) = `r round((ex2_ANOVAtableT9.10[1,5]),2)`, *p* < `r format.pval((ex2_ANOVAtableT9.10[1,6]), digits = 2, eps = .001)`), and adults (*F*(`r ex2_ANOVAtableTAdult[1,3]`, `r ex2_ANOVAtableTAdult[1,4]`) = `r round((ex2_ANOVAtableTAdult[1,5]),2)`, *p* `r format.pval((ex2_ANOVAtableTAdult[1,6]), digits = 2, eps = .001)`) selected more extreme exemplars when asked to select the ideal than when asked to pick the most representative. Also as in Study 1, induction items revealed that generalizations based on the extreme instead of average exemplar declined with age (Wald $X^2$(`r ex2_ANOVAtableD[2,2]`) = `r round((ex2_ANOVAtableD[2,1]),2)`, *p* = `r format.pval((ex2_ANOVAtableD[2,3]), digits = 2, eps = .001)`; ages 5-6: *M* = `r round((meansAgeD_ex2[1,2]),2)`, 95% CI [`r round((meansAgeD_ex2[1,4]),2)`, `r round((meansAgeD_ex2[1,5]),2)`]; ages 7-8: *M* = `r round((meansAgeD_ex2[2,2]),2)`, 95% CI [`r round((meansAgeD_ex2[2,4]),2)`, `r round((meansAgeD_ex2[2,5]),2)`]; ages 9-10: *M* = `r round((meansAgeD_ex2[3,2]),2)`, 95% CI [`r round((meansAgeD_ex2[3,4]),2)`, `r round((meansAgeD_ex2[3,5]),2)`]; adults: *M* = `r round((meansAgeD_ex2[4,2]),2)`, 95% CI [`r round((meansAgeD_ex2[4,4]),2)`, `r round((meansAgeD_ex2[4,5]),2)`]). Participants' choices for ideal exemplars were again less extreme when the property was negative from a human perspective (*F*(`r ex2_ANOVAtableT_Ideals[2,3]`, `r ex2_ANOVAtableT_Ideals[2,4]`) = `r round((ex2_ANOVAtableT_Ideals[2,5]),2)`, *p* = `r format.pval((ex2_ANOVAtableT_Ideals[2,6]), digits = 2, eps = .001)`), with no valence effects on representativeness (*F*(`r ex2_ANOVAtableT_Representative[2,3]`, `r ex2_ANOVAtableT_Ideals[2,4]`) = `r round((ex2_ANOVAtableT_Representative[2,5]),2)`, *p* = `r format.pval((ex2_ANOVAtableT_Representative[2,6]), digits = 2, eps = .001)`) or induction (Wald $X^2$(`r ex2_ANOVAtableD[4,2]`) = `r round((ex2_ANOVAtableD[4,1]),2)`, *p* = `r format.pval((ex2_ANOVAtableD[4,3]), digits = 2, eps = .001)`).

**Study 3**
**Participants**
In Study 3, children were ages 5-6 (*N* = `r nrow(ex3Wide_Child_5.6)`, *M* age = `r round(mean(ex3Wide_Child_5.6$ExactAge, na.rm = TRUE), 2)`; `r sum(ex3Wide_Child_5.6$Gender == "Male", na.rm = TRUE)` male) and 7-8 year olds (*N* =  `r nrow(ex3Wide_Child_7.8)` , *M* age = `r round(mean(ex3Wide_Child_7.8$ExactAge, na.rm = TRUE), 2)`; `r sum(ex3Wide_Child_7.8$Gender == "Male", na.rm = TRUE)` male). Adults in Study 3 (*N* = `r nrow(ex3Wide_Adult)`, *M* age = `r round(mean(ex3Wide_Adult$ExactAge, na.rm = TRUE), 2)`; `r sum(ex3Wide_Adult$Gender == "Male", na.rm = TRUE)` male) were recruited with Amazon Mechanical Turk and tested using Testable.org.

**Results**
In the baseline condition, when participants were introduced to the five exemplars of the novel categories that varied perceptually (e.g., “tulvers” varying in snout length) but were not told about functional consequences or contrastive value of the property, participants of each age selected exemplars around the middle of the scale (i.e., 3) as both most representative (5-6 year olds: *M* = `r round((meansCellT_ex3[1,4]),2)`, 95% CI [`r round((meansCellT_ex3[1,6]),2)`, `r round((meansCellT_ex3[1,7]),2)`]; 7-8 year olds: *M* = `r round((meansCellT_ex3[2,4]),2)`, 95% CI [`r round((meansCellT_ex3[2,6]),2)`, `r round((meansCellT_ex3[2,7]),2)`]; adults: *M* = `r round((meansCellT_ex3[3,4]),2)`, 95% CI [`r round((meansCellT_ex3[3,6]),2)`, `r round((meansCellT_ex3[3,7]),2)`]) and generalizable (5-6 year olds: *M* = `r round((meansCellD_ex3[1,4]),2)`, 95% CI [`r round((meansCellD_ex3[1,6]),2)`, `r round((meansCellD_ex3[1,7]),2)`]; 7-8 year olds: *M* = `r round((meansCellD_ex3[2,4]),2)`, 95% CI [`r round((meansCellD_ex3[2,6]),2)`, `r round((meansCellD_ex3[2,7]),2)`]; adults: *M* = `r round((meansCellD_ex3[3,4]),2)`, 95% CI [`r round((meansCellD_ex3[3,6]),2)`, `r round((meansCellD_ex3[3,7]),2)`]). This pattern indicates that children’s responses in Studies 1-2 cannot be explained by a low-level bias to always pick exemplars on one side of the scale, or to always select exemplars that had “more” of something, paving the way to consider the role of more theoretically meaningful mechanisms.

Consistent with the possibility that participants judge extreme exemplars as representative because they more closely approximate ideal forms, participants did so more often when they received functional information about the property (indicating it was normatively tied to category membership) than when they did not receive this information (and instead had access only to perceptual variation; main effect of functional information (*F*(`r ex3_ANOVAtableT[3,3]`, `r ex3_ANOVAtableT[3,4]`) = `r round((ex3_ANOVAtableT[3,5]),2)`, *p* `r format.pval((ex3_ANOVAtableT[3,6]), digits = 2, eps = .001)`). Although the effect of functional information varied by age group (*F*(`r ex3_ANOVAtableT[5,3]`, `r ex3_ANOVAtableT[5,4]`) = `r round((ex3_ANOVAtableT[5,5]),2)`, *p* `r format.pval((ex3_ANOVAtableT[5,6]), digits = 2, eps = .001)`), at each age participants picked more extreme values when they were given functional information than when they were not (Figure 4). The interaction reflected that this effect was more pronounced among children ages 7-8 than either other group, but the effect of functional information was similar and reliable at each age considered separately (ages 5-6: *F*(`r ex3_ANOVAtableT5.6[2,3]`, `r ex3_ANOVAtableT5.6[2,4]`) = `r round((ex3_ANOVAtableT5.6[2,5]),2)`, *p* `r format.pval((ex3_ANOVAtableT5.6[2,6]), digits = 2, eps = .001)`; ages 7-8: *F*(`r ex3_ANOVAtableT7.8[2,3]`, `r ex3_ANOVAtableT7.8[2,4]`) = `r round((ex3_ANOVAtableT7.8[2,5]),2)`, *p* `r format.pval((ex3_ANOVAtableT7.8[2,6]), digits = 2, eps = .001)`; adults: *F*(`r ex3_ANOVAtableTAdult[2,3]`, `r ex3_ANOVAtableTAdult[2,4]`) = `r round((ex3_ANOVAtableTAdult[2,5]),2)`, *p`r format.pval((ex3_ANOVAtableTAdult[2,6]), digits = 2, eps = .001)`).

Participants were also more likely to select extreme exemplars as providing the most generalizable information (revealed by their decisions of which exemplars to examine to learn about the category) when they received functional information about the properties than otherwise (main effect of functional information, *F*(`r ex3_ANOVAtableD[3,3]`, `r ex3_ANOVAtableD[3,4]`) = `r round((ex3_ANOVAtableD[3,5]),2)`, *p* `r format.pval((ex3_ANOVAtableD[3,6]), digits = 2, eps = .001)`). Although children ages 7-8 selected more extreme exemplars overall (*M* = `r round((meansAgeD_ex3[2,2]),2)`, 95% CI [`r round((meansAgeD_ex3[2,4]),2)`, `r round((meansAgeD_ex3[2,5]),2)`]) than 5-6 year olds (*M* = `r round((meansAgeD_ex3[1,2]),2)`, 95% CI [`r round((meansAgeD_ex3[1,4]),2)`, `r round((meansAgeD_ex3[1,5]),2)`]) or adults (*M* = `r round((meansAgeD_ex3[3,2]),2)`, 95% CI [`r round((meansAgeD_ex3[3,4]),2)`, `r round((meansAgeD_ex3[3,5]),2)`]), the effect of functional information on sample selection did not vary by age group (*F*(`r ex3_ANOVAtableD[5,3]`, `r ex3_ANOVAtableD[5,4]`) = `r round((ex3_ANOVAtableD[5,5]),2)`, *p* = `r format.pval((ex3_ANOVAtableD[5,6]), digits = 2, eps = .001)`) and was reliable at each age examined separately (ages 5-6: *F*(`r ex3_ANOVAtableD5.6[2,3]`, `r ex3_ANOVAtableD5.6[2,4]`) = `r round((ex3_ANOVAtableD5.6[2,5]),2)`, *p* = `r format.pval((ex3_ANOVAtableD5.6[2,6]), digits = 2, eps = .001)`; ages 7-8: *F*(`r ex3_ANOVAtableD7.8[2,3]`, `r ex3_ANOVAtableD7.8[2,4]`) = `r round((ex3_ANOVAtableD7.8[2,5]),2)`, *p* `r format.pval((ex3_ANOVAtableD7.8[2,6]), digits = 2, eps = .001)`; adults: *F*(`r ex3_ANOVAtableDAdult[2,3]`, `r ex3_ANOVAtableDAdult[2,4]`) = `r round((ex3_ANOVAtableDAdult[2,5]),2)`, *p* `r format.pval((ex3_ANOVAtableDAdult[2,6]), digits = 2, eps = .001)`).

Inconsistent with the possibility that people view extreme exemplars as representative because they are more helpful for differentiating categories from one another, participants were not more likely to select extreme exemplars that would be helpful for differentiation in the presence of contrast categories (in fact they were slightly *less* likely to do so; *F*(`r ex3_ANOVAtableT[2,3]`, `r ex3_ANOVAtableT[2,4]`) = `r round((ex3_ANOVAtableT[2,5]),2)`, *p* = `r format.pval((ex3_ANOVAtableT[2,6]), digits = 2, eps = .001)`; contrast: *M* = `r round((meansContrastT[2,2]),2)`, 95% CI [`r round((meansContrastT[2,4]),2)`, `r round((meansContrastT[2,5]),2)`]; no contrast: *M* = `r round((meansContrastT[1,2]),2)`, 95% CI [`r round((meansContrastT[1,4]),2)`, `r round((meansContrastT[1,5]),2)`]). There were no effects of the presence of contrast categories on sample selection responses.
```{r}
Figure4
```
**Figure 4.** Average exemplar selected, by age group and functional information, collapsed across contrast category (N = 296). Higher values represent more functionally valuable exemplars. Exemplar 3 represents the average. Large shapes are means by age group and condition, small lines are responses to individual trials. Facets show representativeness (e.g., "the real, tulver-y tulver") and sample selection responses (e.g., "which would you look at to learn about tulvers?")
```{r, include=FALSE}
#Supplementary analyses

#Preliminary analyses testing for effects of participant background variables from science and nature questionnaire, and interactions with key independent variables.

#Make subset of participants including only those with nature and science background info
ex1_Nature <- subset(ex1Wide, !is.na(ex1Wide$ActivitiesComp))

#Choice
ex1_choice_Nature <- lm(choice_av ~ AgeGroup*Condition + Pets + Camping + Hiking + Fishing + Gardening + CarePets + CareFarmAnimals + Zoos + NatureInterest + NatureBooks + AMNHVisits + DRVisits, ex1_Nature)
ex1_choice_NatureS<-summary(ex1_choice_Nature)
ex1_tableT_Nature<-as.data.frame(coef(ex1_choice_NatureS))
ex1_tableT_Nature
ex1_ANOVAtableT_Nature<-as.data.frame(Anova(ex1_choice_Nature, type = "III", singular.ok = TRUE))
ex1_ANOVAtableT_Nature

ex1_choice_Nature2 <- lm(choice_av ~ AgeGroup*Condition + ActivitiesComp + NatureComp, ex1_Nature)
ex1_choice_Nature2S<-summary(ex1_choice_Nature2)
ex1_tableT_Nature2<-as.data.frame(coef(ex1_choice_Nature2S))
ex1_tableT_Nature2
ex1_ANOVAtableT_Nature2<-as.data.frame(Anova(ex1_choice_Nature2, type = "III", singular.ok = TRUE))
ex1_ANOVAtableT_Nature2
```

```{r, include = FALSE}
#Induction
ex1_induction_Nature <- lm(induction_av ~ AgeGroup*Condition + Pets + Camping + Hiking + Fishing + Gardening + CarePets + CareFarmAnimals + Zoos + NatureInterest + NatureBooks + AMNHVisits + DRVisits, ex1_Nature)
ex1_induction_NatureS<-summary(ex1_induction_Nature)
ex1_tableT_Nature<-as.data.frame(coef(ex1_induction_NatureS))
ex1_ANOVAtableT_Nature<-as.data.frame(Anova(ex1_induction_Nature, type = "III", singular.ok = TRUE))
ex1_ANOVAtableT_Nature

ex1_induction_Nature2 <- lm(induction_av ~ AgeGroup*Condition + ActivitiesComp + NatureComp, ex1_Nature)
ex1_induction_Nature2S<-summary(ex1_induction_Nature2)
ex1_tableT_Nature2<-as.data.frame(coef(ex1_induction_Nature2S))
ex1_ANOVAtableT_Nature2<-as.data.frame(Anova(ex1_induction_Nature2, type = "III", singular.ok = TRUE))
ex1_ANOVAtableT_Nature2
```

```{r, include = FALSE}
#Study 2

#Subset of participants with nature and science background info
ex2_Nature <- subset(ex2Wide, !is.na(ex2Wide$ActivitiesComp))
View(ex2Wide)

#Choice
ex2_choice_Nature <- lm(choice_av ~ AgeGroup*Condition + Pets + Camping + Hiking + Fishing + Gardening + CarePets + CareFarmAnimals + Zoos + NatureInterest + NatureBooks + AMNHVisits + DRVisits, ex2_Nature)
ex2_choice_NatureS<-summary(ex2_choice_Nature)
ex2_tableT_Nature<-as.data.frame(coef(ex2_choice_NatureS))
ex2_tableT_Nature
ex2_ANOVAtableT_Nature<-as.data.frame(Anova(ex2_choice_Nature, type = "III", singular.ok = TRUE))
ex2_ANOVAtableT_Nature

ex2_choice_Nature2 <- lm(choice_av ~ AgeGroup*Condition + ActivitiesComp + NatureComp, ex2_Nature)
ex2_choice_Nature2S<-summary(ex2_choice_Nature2)
ex2_tableT_Nature2<-as.data.frame(coef(ex2_choice_Nature2S))
ex2_tableT_Nature2
ex2_ANOVAtableT_Nature2<-as.data.frame(Anova(ex2_choice_Nature2, type = "III", singular.ok = TRUE))
ex2_ANOVAtableT_Nature2
View(ex1Wide)
```

```{r, include = FALSE}
#Induction
ex2_induction_Nature <- lm(induction_av ~ AgeGroup*Condition + Pets + Camping + Hiking + Fishing + Gardening + CarePets + CareFarmAnimals + Zoos + NatureInterest + NatureBooks + AMNHVisits + DRVisits, ex2_Nature)
ex2_induction_NatureS<-summary(ex2_induction_Nature)
ex2_tableT_Nature<-as.data.frame(coef(ex2_induction_NatureS))
ex2_ANOVAtableT_Nature<-as.data.frame(Anova(ex2_induction_Nature, type = "III", singular.ok = TRUE))
ex2_ANOVAtableT_Nature

ex2_induction_Nature2 <- lm(induction_av ~ AgeGroup*Condition + ActivitiesComp + NatureComp, ex2_Nature)
ex2_induction_Nature2S<-summary(ex2_induction_Nature2)
ex2_tableT_Nature2<-as.data.frame(coef(ex2_induction_Nature2S))
ex2_ANOVAtableT_Nature2<-as.data.frame(Anova(ex2_induction_Nature2, type = "III", singular.ok = TRUE))
ex2_ANOVAtableT_Nature2
```

```{r, include = FALSE}
# Study 3

#Subset of participants with nature and science background info
ex3_Nature <- subset(ex3Wide, !is.na(ex3Wide$ActivitiesComp))

#Preliminary analysis including participant background variables
#Choice
ex3_choice_Nature <- lm(choice_av ~ AgeGroup*Functional*Contrast + Pets + Camping + Hiking + Fishing + Gardening + CarePets + CareFarmAnimals + Zoos + NatureInterest + NatureBooks + AMNHVisits + DRVisits, ex3_Nature)
ex3_choice_NatureS<-summary(ex3_choice_Nature)
ex3_tableT_Nature<-as.data.frame(coef(ex3_choice_NatureS))
ex3_tableT_Nature
ex3_ANOVAtableT_Nature<-as.data.frame(Anova(ex3_choice_Nature, type = "III", singular.ok = TRUE))
ex3_ANOVAtableT_Nature

ex3_choice_Nature2 <- lm(choice_av ~ AgeGroup*Functional*Contrast + ActivitiesComp + NatureComp, ex3_Nature)
ex3_choice_Nature2S<-summary(ex3_choice_Nature2)
ex3_tableT_Nature2<-as.data.frame(coef(ex3_choice_Nature2S))
ex3_tableT_Nature2
ex3_ANOVAtableT_Nature2<-as.data.frame(Anova(ex3_choice_Nature2, type = "III", singular.ok = TRUE))
ex3_ANOVAtableT_Nature2
```

```{r, include = FALSE}
#Induction
ex3_induction_Nature <- lm(induction_av ~ AgeGroup*Functional*Contrast + Pets + Camping + Hiking + Fishing + Gardening + CarePets + CareFarmAnimals + Zoos + NatureInterest + NatureBooks + AMNHVisits + DRVisits, ex3_Nature)
ex3_induction_NatureS<-summary(ex3_induction_Nature)
ex3_tableT_Nature<-as.data.frame(coef(ex3_induction_NatureS))
ex3_ANOVAtableT_Nature<-as.data.frame(Anova(ex3_induction_Nature, type = "III", singular.ok = TRUE))
ex3_ANOVAtableT_Nature

ex3_induction_Nature2 <- lm(induction_av ~ AgeGroup*Functional*Contrast + ActivitiesComp + NatureComp, ex3_Nature)
ex3_induction_Nature2S<-summary(ex3_induction_Nature2)
ex3_tableT_Nature2<-as.data.frame(coef(ex3_induction_Nature2S))
ex3_ANOVAtableT_Nature2<-as.data.frame(Anova(ex3_induction_Nature2, type = "III", singular.ok = TRUE))
ex3_ANOVAtableT_Nature2
```
#Supplementary Online Materials

**Science and Nature Questionnaire**
Children in Studies 1-3 were recruited and tested in a major cultural institution (The American Museum of Natural History). Of participants who provided location information, `r  sum(Demo_Child$Community == "Urban", na.rm = TRUE)` were from urban environments, `r  sum(Demo_Child$Community == "Suburban", na.rm = TRUE)` from suburban environments, and `r  sum(Demo_Child$Community == "Rural", na.rm = TRUE)` from rural environments. In order to test for possible effects of children's experiences on our key variables, we gathered information on children's background experiences with science and nature in the form of a questionnaire completed by parents, as has been used in previous work at this institution (Rhodes & Liebenson, 2015). The complete Science and Nature Background Questionnaire used in Studies 1-3 is available at https://osf.io/v5pcs/.

Of the `r nrow(ex1Wide_Child)` child participants in Study 1, `r  nrow(ex1_Nature)` families chose to provide information about background experiences with science and nature; `r  nrow(ex2_Nature)` of the total `r nrow(ex2Wide_Child)` children in Study 2 and `r  nrow(ex3_Nature)` of the `r nrow(ex3Wide_Child)` children in Study 3 provided this information. In preliminary analyses for each study, we tested whether children's representativeness judgments and inductive inferences were significantly affected by background variables including the child's home community, the presence of pets in the home, each item on the 6-point "activities in the past year" scale (and a composite of these questions), each item on the 4-point "interest in nature" scale (and a composite of these questions), number of visits to the museum, and number of visits to the interactive Discovery Room portion of the museum where the study took place. None of the background variables significantly predicted responses or interacted with any of our key independent variables, therefore we did not consider these participant variables further.
```{r, include = FALSE}
#Supplementary analysis for Study 3, with direction included in the model. Direction was coded as "most-least" = animal items for which the progression was most to least (with the "most" of the property as functionally better), and "least-most" = animal items for which the progression was least to most (with the "least" of the property as functionally better). The goal of these analyses is to capture whether the effect of functional information can be explained by a low-level preference to choose exemplars with the "most" of a varied property.

#Representativeness
Representative_dir <- lmer(cbind(representative) ~ AgeGroup*Contrast*Functional*direction + (1|SID) + (1|item), data = ex3)
ex3_ANOVAtableT_dir<-as.data.frame(anova(Representative_dir, ddf = "Kenward-Roger", type = 3))
ex3_ANOVAtableT_dir

meansAgeDirT<-effect("AgeGroup*direction", Representative_dir)
meansAgeDirT<-as.data.frame(meansAgeDirT)
meansAgeDirT

meansConDirT<-effect("Contrast*direction", Representative_dir)
meansConDirT<-as.data.frame(meansConDirT)
meansConDirT
```

```{r, include = FALSE}
Discover_dir <- lmer(cbind(induction) ~ AgeGroup*Contrast*Functional*direction + (1|SID) + (1|item), data = ex3)
ex3_ANOVAtableD_dir<-as.data.frame(anova(Discover_dir, ddf = "Kenward-Roger", type = 3))
ex3_ANOVAtableD_dir

meansAgeDirD<-effect("AgeGroup*direction", Discover_dir)
meansAgeDirD<-as.data.frame(meansAgeDirD)
meansAgeDirD
```
**Study 3 Supplementary Analyses**
To address possible low-level mechanisms that could explain the pattern of results in Studies 1-2, in Study 3 we varied the nature of the properties shown in the novel animal scales, so that having *more* of a target property facilitated the key behavior (e.g., longer snouts helped dig for food) for half of the items, whereas having *less* did so for the other half (e.g., smaller ears were better for hiding from predators). We ran preliminary analyses including the direction of the scale in the models for representativeness and sample selection choices in order ensure that any effects of functional and contrastive value could not be explained by these low-level perceptual mechanisms. If younger children selected more extreme exemplars in Studies 1-2 only because of these perceptual cues, and selected exemplars with the "most" of a perceptually salient property as representative and informative regardless of category knowledge, then the direction of the scale alone would predict responses on these measures in Study 3.

The effect of direction across trials (animal kinds with a least-to-most progression versus those with a most-to-least progression) varied with age (*F*(`r ex3_ANOVAtableT_dir[8,3]`, `r ex3_ANOVAtableT_dir[8,4]`) = `r round((ex3_ANOVAtableT_dir[8,5]),2)`, *p* = `r format.pval((ex3_ANOVAtableT_dir[8,6]), digits = 2, eps = .001)`; subsumed main effect of age group, *F*(`r ex3_ANOVAtableT_dir[1,3]`, `r ex3_ANOVAtableT_dir[1,4]`) = `r round((ex3_ANOVAtableT_dir[1,5]),2)`, *p* = `r format.pval((ex3_ANOVAtableT_dir[1,6]), digits = 2, eps = .001)`; subsumed main effect of direction, *F*(`r ex3_ANOVAtableT_dir[4,3]`, `r ex3_ANOVAtableT_dir[4,4]`) = `r round((ex3_ANOVAtableT_dir[4,5]),2)`, *p* = `r format.pval((ex3_ANOVAtableT_dir[4,6]), digits = 2, eps = .001)`). Participants of all ages chose more extreme exemplars overall on trials with a least-to-most progression, but this was most pronounced in younger children (least-to-most, ages 5-6: *M* = `r round((meansAgeDirT[1,3]),2)`,  CI [`r round((meansAgeDirT[1,5]),2)`, `r round((meansAgeDirT[1,6]),2)`]; ages 7-8:  *M* = `r round((meansAgeDirT[2,3]),2)`,  CI [`r round((meansAgeDirT[2,5]),2)`, `r round((meansAgeDirT[2,6]),2)`], adults: *M* = `r round((meansAgeDirT[3,3]),2)`,  CI [`r round((meansAgeDirT[3,5]),2)`, `r round((meansAgeDirT[3,6]),2)`], most-to-least, ages 5-6: *M* = `r round((meansAgeDirT[4,3]),2)`,  CI [`r round((meansAgeDirT[4,5]),2)`, `r round((meansAgeDirT[4,6]),2)`]; ages 7-8: *M* = `r round((meansAgeDirT[5,3]),2)`,  CI [`r round((meansAgeDirT[5,5]),2)`, `r round((meansAgeDirT[5,6]),2)`]; adults: *M* = `r round((meansAgeDirT[6,3]),2)`,  CI [`r round((meansAgeDirT[6,5]),2)`, `r round((meansAgeDirT[6,6]),2)`]). Direction of the scale did not interact with functional information (*F*(`r ex3_ANOVAtableT_dir[10,3]`) = `r round((ex3_ANOVAtableT_dir[10,5]),2)`, *p* = `r format.pval((ex3_ANOVAtableT_dir[10,6]), digits = 2, eps = .001)`).

Similar patterns emerged for sample selection choices (age group x direction interaction, *F*(`r ex3_ANOVAtableD_dir[8,3]`, `r ex3_ANOVAtableD_dir[8,4]`) = `r round((ex3_ANOVAtableD_dir[8,5]),2)`, *p* = `r format.pval((ex3_ANOVAtableD_dir[8,6]), digits = 2, eps = .001)`; subsumed main effects of age group, *F*(`r ex3_ANOVAtableD_dir[1,3]`, `r ex3_ANOVAtableD_dir[1,4]`) = `r round((ex3_ANOVAtableD_dir[1,5]),2)`, *p* = `r format.pval((ex3_ANOVAtableD_dir[1,6]), digits = 2, eps = .001)`, and direction, *F*(`r ex3_ANOVAtableD_dir[4,3]`, `r ex3_ANOVAtableT_dir[4,4]`) = `r round((ex3_ANOVAtableD_dir[4,5]),2)`, *p* = `r format.pval((ex3_ANOVAtableD_dir[4,6]), digits = 2, eps = .001)`). Participants of all ages chose to learn from relatively more extreme exemplars on animal trials with a least-to-most direction, but these effects were most pronounced in younger children (least-to-most, ages 5-6: *M* = `r round((meansAgeDirD[1,3]),2)`,  CI [`r round((meansAgeDirD[1,5]),2)`, `r round((meansAgeDirD[1,6]),2)`]; ages 7-8: *M* = `r round((meansAgeDirD[2,3]),2)`,  CI [`r round((meansAgeDirD[2,5]),2)`, `r round((meansAgeDirD[2,6]),2)`]; adults: *M* = `r round((meansAgeDirD[3,3]),2)`,  CI [`r round((meansAgeDirD[3,5]),2)`, `r round((meansAgeDirD[3,6]),2)`]; trials with a most-to-least progression, ages 5-6: *M* = `r round((meansAgeDirD[4,3]),2)`,  CI [`r round((meansAgeDirD[4,5]),2)`, `r round((meansAgeDirD[4,6]),2)`]; ages 7-8: *M* = `r round((meansAgeDirD[5,3]),2)`,  CI [`r round((meansAgeDirD[5,5]),2)`, `r round((meansAgeDirD[5,6]),2)`]; adults: *M* = `r round((meansAgeDirD[6,3]),2)`,  CI [`r round((meansAgeDirD[6,5]),2)`, `r round((meansAgeDirD[6,6]),2)`]).

While these analyses suggest that younger children did have an overall bias to pick exemplars that had more (rather than less) of a perceptual property, this tendency does not account for the full pattern of findings obtained in Study 3 or alter our interpretation of those results. As shown in Figure S1 and S2, functional information reliably led participants of all ages to select more extreme exemplars as both more representative (Figure S1) and informative (Figure S2) on all items—including both those where having more of a property was more functional and those where having less of a property was more functional.

This analysis also revealed that Direction information interacted with whether children were told about a Contrast category or not, but not in a manner indicating that contrast categories facilitate selections of extreme exemplars. As reported in the main text, participants did not choose more constrastive exemplars in the presence of relevant contrast categories (in fact, they chose slightly *less* contrastive ones, main effect of contrast category: *F*(`r ex3_ANOVAtableT[2,3]`, `r ex3_ANOVAtableT_dir[2,4]`) = `r round((ex3_ANOVAtableT[2,5]),2)`, *p* = `r format.pval((ex3_ANOVAtableT[2,6]), digits = 2, eps = .001)`). This supplemental analysis revealed that this pattern emerged only on trials with a most-to-least progression--on these trials, participants chose exemplars that were *more* similar to the contrast category (meaning that they also displayed *more* of the varied property; *M* = `r round((meansAgeDirT[4,3]),2)`,  CI [`r round((meansAgeDirT[4,5]),2)`, `r round((meansAgeDirT[4,6]),2)`]; no contrast: *M* = `r round((meansAgeDirT[3,3]),2)`,  CI [`r round((meansAgeDirT[3,5]),2)`, `r round((meansAgeDirT[3,6]),2)`]). On trials with a least-to-most progression, however, participants responded similarly regardless of the presence or absence of a relevant contrast category (contrast: *M* = `r round((meansConDirT[2,3]),2)`,  CI [`r round((meansConDirT[2,5]),2)`, `r round((meansConDirT[2,6]),2)`]; no contrast: *M* = `r round((meansConDirT[1,3]),2)`,  CI [`r round((meansConDirT[1,5]),2)`, `r round((meansConDirT[1,6]),2)`]; contrast x direction interaction: *F*(`r ex3_ANOVAtableT_dir[9,3]`, `r ex3_ANOVAtableT_dir[9,4]`) = `r round((ex3_ANOVAtableT_dir[9,5]),2)`, *p* = `r format.pval((ex3_ANOVAtableT_dir[9,6]), digits = 2, eps = .001)`).
```{r, include = FALSE}
#Figure for items in Study 3, No Contrast conditions only, with panels for each item and image inserted, and means by age group and functional.

#Give labels to items so that they can be associated with the images
NoContrast$item<-as.factor(NoContrast$item)
NoContrast$item<-recode_factor(NoContrast$item, "1" = "Tulver", "2" = "Virdex", "3"  = "Bosa", "4" = "Sibu", "5" = "Akar", "6" = "Lorp", "7" = "Kita", "8" = "Upos")

#reorder functional info so it graphs in the same order as previous graphs
NoContrast$Functional <- factor(NoContrast$Functional, levels = c("Functional", "Nonfunctional"))

meansItem_rep <- NoContrast %>%
  group_by(AgeGroup, Functional, item) %>%
       summarise(representative=mean(representative))
```

```{r, echo = FALSE, warning = FALSE}
#Create Figure 5--representativeness
FigureS1<-ggplot() +
  geom_hline(mapping = NULL, data = NULL, colour = "#F9610A", yintercept = 3, na.rm = FALSE,
  show.legend = NA) +
  geom_point(data = meansItem_rep, aes(x = AgeGroup, y = representative, colour = Functional, shape = Functional), stat = "identity", size = 2) +
    scale_x_discrete(limits=c("Child_5.6", "Child_7.8", "Adult"), labels = c("Child_5.6" = "5-6 yrs","Child_7.8" = "7-8 yrs", "Adult" = "Adults")) +
  ylim(1,5) +
  geom_jitter(data = NoContrast, aes(x = AgeGroup, y = representative, colour = Functional), width = .30, height = .00, alpha = .5, shape = 124, size = 1) +
  ggtitle("Representative choices by item") +
  facet_wrap(~item, ncol=4) +
  theme_few() +
  theme(axis.text=element_text(size=7),
        axis.title=element_text(size=7)) +
  theme(legend.position="bottom", legend.direction="horizontal") +
  theme(legend.title = element_blank()) +
  scale_shape_manual(values=c(17, 19)) +
  scale_color_manual(values=c("#4C9900", "#0066CC"), guide = guide_legend(reverse = FALSE)) +
  labs(x="Age Group", y="Average exemplar selected")

FigureS1
```

**Figure S1.** Average exemplar selected on the representative task (e.g., "Which is the *real, tulver-y tulver*?"), by age group and functional information, no contrast conditions only. Higher values represent more functionally valuable exemplars. Exemplar 3 represents the average. Large dots are means by age group, small lines are individual participant selections. Facets show responses by item. Participants selected more extreme exemplars when given functional information at all age groups, on every item.
```{r, include = FALSE}
#Discover (induction)
meansItem_dis <- NoContrast %>%
  group_by(AgeGroup, Functional, item) %>%
       summarise(induction=mean(induction))
```

```{r, echo = FALSE, warning = FALSE}
#Create Figure 6--discover (sample selection)
FigureS2<-ggplot() +
  geom_hline(mapping = NULL, data = NULL, colour = "#F9610A", yintercept = 3, na.rm = FALSE,
  show.legend = NA) +
  geom_point(data = meansItem_dis, aes(x = AgeGroup, y = induction, colour = Functional, shape = Functional), stat = "identity", size = 2) +
  scale_x_discrete(limits=c("Child_5.6", "Child_7.8", "Adult"), labels = c("Child_5.6" = "5-6","Child_7.8" = "7-8", "Adult" = "Adults")) +
  ylim(1,5) +
  geom_jitter(data = NoContrast, aes(x = AgeGroup, y = induction, colour = Functional), width = .30, height = .00, alpha = .5, shape = 124, size = 1) +
  ggtitle("Sample selection choices by item") +
  facet_wrap(~item, ncol=4) +
  theme_few() +
  theme(axis.text=element_text(size=7),
        axis.title=element_text(size=7)) +
  theme(legend.position="bottom", legend.direction="horizontal") +
  theme(legend.title = element_blank()) +
  scale_shape_manual(values=c(17, 19)) +
  scale_color_manual(values=c("#4C9900", "#0066CC"), guide = guide_legend(reverse = FALSE)) +
    labs(x="Age Group", y=NULL)

FigureS2
```

**Figure S2.** Average exemplar selected on the sample selection task (e.g., "Which would you look at to learn the most about tulvers?"), by age group and functional information, no contrast conditions only. Higher values represent more functionally valuable exemplars. Exemplar 3 represents the average. Large dots are means by Age Group, small lines are individual participant selections. Facets show responses by item. Participants selected more extreme exemplars when given functional information on almost all items. Note that on the "Sibu" animal trial, participants saw five novel animals varying in mane fluffiness; in functional information conditions, participants were told that having smoother (i.e., less fluffy) manes was more functionally ideal for running fast to catch prey. This explanation may have been too complicated for children and could have conflicted with an existing bias to view fluffier manes as more ideal (e.g., having a fluffy mane is a characteristic property of lions).